---
title: "Selection"
output: 
  html_document:
    css: css\style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
library(flextable)
```

This section is designed to help you understand the concepts behind selection.

## No Selection

Suppose you have a population with genotypic frequency given by the picture and table below.  

 <div class="row" style= "display: flex; align-items: center; justify-content: center;">
  <div class="column">
  
```{r}
MM <- 20
MN <- 120
NN <- 60
Total = MM+MN+NN
df <- data.frame(Genotype = c("#Individuals", "Proportion"), MM = c(MM, MM/Total), MN = c(MN, MN/Total), NN = c(NN, NN/Total), Total = c(Total, 1))
df %>% regulartable() %>% autofit() %>% fontsize(size = 14, part = "all")
```

</div>
  <div class="column">
  ```{r, fig.align="center"}
plot(c(0, 1), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)
rect(0, 0, 0.1, 1, density = 50, col = "Blue")
rect(0.1, 0, 0.7, 1, density = 50, col = "Purple")
rect(0.7, 0, 1, 1, density = 50, col = "Red")
text("MM", x = 0.05, y = 0.5, cex = 2)
text("MN", x = 0.4, y = 0.5, cex = 2)
text("NN", x = 0.85, y = 0.5, cex = 2)
```
</div>
</div> 


Making the assumptions required for the HW law, one thing implied is that there is no selection being done.  That is, all genotypes are equally fit (MM, MN, and NN are not favored in any way).  Under all the HW assumptions, in a single generation this population should reach HW equilibrium given by the table below.  

<br>

```{r}
MM <- 20
MN <- 120
NN <- 60
Total <- MM+MN+NN
p <- (MM+0.5*MN)/Total 
sq <- data.frame(First = c("", paste("M, p = ", round(p,4)), paste("N, 1-p = ", round(1-p,4))), Second = c(paste("M, p = ", round(p,4)), paste("p^2 = ",round(p^2,4)), paste("p*(1-p) = ", round(p*(1-p),4))), Third = c(paste("N, 1-p = ", round(1-p,4)), paste("p*(1-p) = ", round(p*(1-p),4)), paste("(1-p)^2 = ", round((1-p)^2,4))))
colnames(sq) <- sq[,1]
sq <- sq[-1,]
rownames(sq) <- NULL
knitr::kable(sq)
```

<br>

<div class = "row">
<div class= "col-sm-4">
Parent generation:  

```{r}
MM <- 20
MN <- 120
NN <- 60
Total <- MM + MN + NN
df <- data.frame(Genotype = c("#Individuals", "Proportion"), MM = c(MM, MM/Total), MN = c(MN, MN/Total), NN = c(NN, NN/Total), Total = c(Total, 1))

df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
```  

```{r, out.width="90%", fig.align="center"}
plot(c(0, 1), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)
rect(0, 0, 0.1, 1, density = 50, col = "Blue")
rect(0.1, 0, 0.7, 1, density = 50, col = "Purple")
rect(0.7, 0, 1, 1, density = 50, col = "Red")
text("MM", x = 0.05, y = 0.5, cex = 2)
text("MN", x = 0.4, y = 0.5, cex = 2)
text("NN", x = 0.85, y = 0.5, cex = 2)
```  

</div> 

<div class = "col-sm-4">

Offspring (F1):  

```{r}
MM <- p^2*Total
MN <- 2*p*(1-p)*Total
NN <- (1-p)^2*Total
Total <- MM + MN + NN
df <- data.frame(Genotype = c("#Individuals", "Proportion"), MM = c(MM, MM/Total), MN = c(MN, MN/Total), NN = c(NN, NN/Total), Total = c(Total, 1))

df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
```

```{r, out.width="90%", fig.align="center"}
plot(c(0, 1), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)
rect(0, 0, p^2, 1, density = 50, col = "Blue")
rect(p^2, 0, 1-(1-p)^2, 1, density = 50, col = "Purple")
rect(1-(1-p)^2, 0, 1, 1, density = 50, col = "Red")
text("MM", x = p^2/2, y = 0.5, cex = 2)
text("MN", x = (p^2+p^2+2*p*(1-p))/2, y = 0.5, cex = 2)
text("NN", x = (1-(1-p)^2+1)/2, y = 0.5, cex = 2)
```

</div>

<div class = "col-sm-4">

Offspring (F2):  

```{r}
MM <- p^2*Total
MN <- 2*p*(1-p)*Total
NN <- (1-p)^2*Total
Total <- MM + MN + NN
df <- data.frame(Genotype = c("#Individuals", "Proportion"), MM = c(MM, MM/Total), MN = c(MN, MN/Total), NN = c(NN, NN/Total), Total = c(Total, 1))
df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
```

```{r, out.width="90%", fig.align="center"}
plot(c(0, 1), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)
rect(0, 0, p^2, 1, density = 50, col = "Blue")
rect(p^2, 0, 1-(1-p)^2, 1, density = 50, col = "Purple")
rect(1-(1-p)^2, 0, 1, 1, density = 50, col = "Red")
text("MM", x = p^2/2, y = 0.5, cex = 2)
text("MN", x = (p^2+p^2+2*p*(1-p))/2, y = 0.5, cex = 2)
text("NN", x = (1-(1-p)^2+1)/2, y = 0.5, cex = 2)
```

</div>
</div>


When there is selection involved, some genotypes are favorable or unfavorable.  Recall weights can be assigned to each genotype: $w_{MM}, w_{MN}, w_{NN}$  

Examples of common types of selection are given below:  

```{r}
selection <- data.frame(Genotypes = c("No Selection", "Directional", "Disruptive", "Stabilizing"), Wmm = c(1, 1, 1, 1), Wmn = c(1, 0.8, 0.3, 1.3), Wnn = c(1, 0.6, 0.7, 0.2))
knitr::kable(selection)
```

Unlike HW equilibrium and no selection, generations will continue to change their genotypic frequencies over time.  The weights can be used to determine the new frequencies.  For this reason we'll just work with the proportions rather than the counts.

Ex: Consider a directional selection having $w_{MM} = 1$, $w_{MN} = 0.5, w_{NN} = 0.25$. 

<div class = "row">
<div class= "col-sm-5">
Current population (Parent generation):  

```{r}
MM <- p^2*Total
MN <- 2*p*(1-p)*Total
NN <- (1-p)^2*Total
Total <- MM + MN + NN
df <- data.frame(Genotype = c( "Proportion"), MM = c( MM/Total), MN = c(MN/Total), NN = c( NN/Total), Total = 1)

df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
```

```{r, out.width="100%", fig.align="center"}
plot(c(0, 1), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)
rect(0, 0, p^2, 1, density = 50, col = "Blue")
rect(p^2, 0, 1-(1-p)^2, 1, density = 50, col = "Purple")
rect(1-(1-p)^2, 0, 1, 1, density = 50, col = "Red")
text("MM", x = p^2/2, y = 0.5, cex = 2)
text("MN", x = (p^2+p^2+2*p*(1-p))/2, y = 0.5, cex = 2)
text("NN", x = (1-(1-p)^2+1)/2, y = 0.5, cex = 2)
```

</div> 

<div class = "col-sm-7">

Offspring (F1):  

```{r}
Wmm <-1
Wmn <- 0.5
Wnn <- 0.25
MM <- p^2*Total
MMw <- Wmm*MM
MN <- 2*p*(1-p)*Total
MNw <- Wmn*MN
NN <- (1-p)^2*Total
NNw <- Wnn*NN
Totalw <- MMw + MNw+ NNw
df <- data.frame(Genotype = c("Proportion", "Updated"), MM = c(paste0("f(MM)*Wmm = \n", round(MM/Total,4),"*", Wmm, " = ", round(MM*Wmm/Total,4)), paste0(round(MM*Wmm/Total,4), "/", round(MM*Wmm/Total + MN*Wmn/Total+NN*Wnn/Total, 4), " = ", round(MMw/Totalw,4))), 
                 MN= c(paste0("f(MN)*Wmn = \n", round(MN/Total,4),"*", Wmn, " = ", round(MN*Wmn/Total,4)), paste0(round(MN*Wmn/Total,4), "/", round(MM*Wmm/Total + MN*Wmn/Total+NN*Wnn/Total, 4), " = ", round(MNw/Totalw,4))),
                 NN = c(paste0("f(NN)*Wnn = \n", round(NN/Total,4),"*", Wnn, " = ", round(NN*Wnn/Total,4)), paste0(round(NN*Wnn/Total,4), "/", round(MM*Wmm/Total + MN*Wmn/Total+NN*Wnn/Total, 4), " = ", round(NNw/Totalw,4))),
                 Total = c(round((MM*Wmm+MN*Wmn+NN*Wnn)/Total,4), 1))
df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
```


```{r, out.width="100%", fig.align="center"}
plot(c(0, 2.2), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)
rect(0, 0, p^2*Wmm, 1, density = 50, col = "Blue")
rect(p^2*Wmm, 0, p^2, 1, density = 50, col = "Gray")
rect(p^2, 0, 2*p*(1-p)*Wmn+p^2, 1, density = 50, col = "Purple")
rect(2*p*(1-p)*Wmn+p^2, 0, 1-(1-p)^2, 1, density = 50, col = "Gray")
rect(1-(1-p)^2, 0, 1-(1-p)^2 + (1-p)^2*Wnn, 1, density = 50, col = "Red")
rect(1-(1-p)^2 + (1-p)^2*Wnn, 0, 1, 1, density = 50, col = "Gray")
text("MM", x = p^2/2, y = 0.5, cex = 2)
text("MN", x = (p^2+p^2+2*p*(1-p))/2, y = 0.5, cex = 2)
text("NN", x = (1-(1-p)^2+1)/2, y = 0.5, cex = 2)

#equal sign
rect(1.05, 0.47, 1.15, 0.49, density = 100, col = "Black")
rect(1.05, 0.51, 1.15, 0.53, density = 100, col = "Black")
#add next standardized part
rect(1.2, 0, 1.2+MMw/Totalw, 1, density = 50, col = "Blue")
rect(1.2+MMw/Totalw, 0, 2.2- NNw/Totalw, 1, density = 50, col = "Purple")
rect(2.2- NNw/Totalw, 0, 2.2, 1, density = 50, col = "Red")
text("MM", x = (1.2+1.2+MMw/Totalw)/2, y = 0.5, cex = 2)
text("MN", x = (1.2+MMw/Totalw + 2.2 - NNw/Totalw)/2, y = 0.5, cex = 2)
text("NN", x = (2.2- NNw/Totalw+2.2)/2, y = 0.5, cex = 2)
```

</div>
</div>

In the plots on the right, the greyed out part represents where the selection is occurring.  Half of the MN and three fourths of the NN genotypes die off.  The area remaining is then restandardized to account for losing this area.  

In the next generation the frequencies would change again.

<div class = "row">
<div class= "col-sm-5">
Current population (F1):  

```{r}
Wmm <-1
Wmn <- 0.5
Wnn <- 0.25
MMp <- MMw/Totalw
MNp <- MNw/Totalw
NNp <- NNw/Totalw

df <- data.frame(Genotype = c("Proportion"), MM = round(MMp, 4), MN = round(MNp, 4), NN = round(NNp, 4), Total = 1)
df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
```


```{r, out.width="100%", fig.align="center"}
plot(c(0, 1), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)

#add next standardized part
rect(0, 0, MMp, 1, density = 50, col = "Blue")
rect(MMp, 0, MMp+MNp, 1, density = 50, col = "Purple")
rect(MMp+MNp, 0, 1, 1, density = 50, col = "Red")
text("MM", x = (MMp)/2, y = 0.5, cex = 2)
text("MN", x = (MMp+MMp+MNp)/2, y = 0.5, cex = 2)
text("NN", x = (MMp+MNp+1)/2, y = 0.5, cex = 2)
```

</div> 

<div class = "col-sm-7">

Offspring (F2):  

```{r}
Wmm <-1
Wmn <- 0.5
Wnn <- 0.25
newTotal <- Wmm*MMp + Wmn*MNp + Wnn*NNp
df <- data.frame(Genotype = c("Proportion", "Updated"), 
                 MM = c(paste0("f(MM)*Wmm = \n", round(MMp,4),"*", Wmm, " = ", round(Wmm*MMp,4)),
                        paste0(round(Wmm*MMp,4), "/", round(newTotal, 4), " = ", round(Wmm*MMp/(newTotal),4))), 
                 MN= c(paste0("f(MN)*Wmn = \n", round(MNp,4),"*", Wmn, " = ", round(Wmn*MNp,4)),
                        paste0(round(Wmn*MMp,4), "/", round(newTotal, 4), " = ", round(Wmn*MNp/(newTotal),4))), 
                 NN = c(paste0("f(NN)*Wnn = \n", round(NNp,4),"*", Wnn, " = ", round(Wnn*NNp,4)),
                        paste0(round(Wnn*NNp,4), "/", round(newTotal, 4), " = ", round(Wnn*NNp/(newTotal),4))), 
                 Total = c(round(newTotal,4), 1))
df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
```

```{r}
parMM <- Wmm*MMp
parMN <- Wmn*MNp
parNN <- Wnn*NNp

neWmm <- Wmm*MMp/(newTotal)
neWmn <- Wmn*MNp/(newTotal)
neWnn <- Wnn*NNp/(newTotal)

plot(c(0, 2.2), c(0, 1), type = "n", xlab = "", ylab = "", axes =FALSE)
rect(0, 0, parMM, 1, density = 50, col = "Blue")
rect(parMM, 0, MMp, 1, density = 50, col = "Gray")
rect(MMp, 0, MMp + parMN, 1, density = 50, col = "Purple")
rect(MMp + parMN, 0, MMp+MNp, 1, density = 50, col = "Gray")
rect(MMp+MNp, 0, MMp+MNp+parNN, 1, density = 50, col = "Red")
rect(MMp+MNp+parNN, 0, 1, 1, density = 50, col = "Gray")
text("MM", x = MMp/2, y = 0.7, cex = 2)
text("MN", x = (MMp+MNp+MMp)/2, y = 0.5, cex = 2)
text("NN", x = (MNp+MMp+1)/2, y = 0.3, cex = 2)

#equal sign
rect(1.05, 0.47, 1.15, 0.49, density = 100, col = "Black")
rect(1.05, 0.51, 1.15, 0.53, density = 100, col = "Black")
#add next standardized part
rect(1.2, 0, 1.2+neWmm, 1, density = 50, col = "Blue")
rect(1.2+neWmm, 0, 1.2+neWmm+neWmn, 1, density = 50, col = "Purple")
rect(1.2+neWmm+neWmn, 0, 2.2, 1, density = 50, col = "Red")
text("MM", x = (1.2+1.2+neWmm)/2, y = 0.7, cex = 2)
text("MN", x = (1.2+neWmm+1.2+neWmm+neWmn)/2, y = 0.5, cex = 2)
text("NN", x = (1.2+neWmm+neWmn+2.2)/2, y = 0.3, cex = 2)
```


</div>
</div>

### Problems  

1. Suppose that your population has genotypic frequencies given by the table below:  

    ```{r}
    MM <- p^2*Total
    MN <- 2*p*(1-p)*Total
    NN <- (1-p)^2*Total
    Total <- MM + MN + NN
    df <- data.frame(Genotype = c( "Proportion"), MM = c( MM/Total), MN = c(MN/Total), NN = c( NN/Total), Total = 1)

    df %>% regulartable() %>% autofit()%>% fontsize(size = 14, part = "all")
    ```
    Find the genotypic frequencies for the F1 generation under the "Stabilizing" selection weights given in the table above.  

2. Use the [application here](https://shiny.stat.ncsu.edu/jbpost2/Selection) to see the behavior of the population over time.  Set the allele frequency to be 0.4 and use the selection weights from question 1.  

    a. How many generations does it take before the NN genotype is essentially gone?  
    
    b. Why does the NN genotype disappear more quickly than the MM genotype?
    

3.  Use your own population genotypic frequencies and selection weights to see they affect genotypic frequencies over time.  

<br>
<br>
<br>
